Relationship Between Common Stream Health Indicators

### 2021-10-29 code/processing_code/Data_Exploration.Rmd
Project Part 3 - CC

#### First, we need to load all the libraries we will need.
```{r}
library(dplyr) 
library(here)
library(tidyverse)
library(tidyr)
library(readr)
library(stringr)
```

@@

## MIDO

@@

#### Biological Vs Conductivity
```{r}
BS_Con <- MIDO %>% 
  ggplot(aes(x=(conductivity.uscm), 
             y=(biological_score))) +
  geom_point(color = "red") +
  geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Turbidity
```{r}
BS_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Nitrate Concentration
```{r}
BS_no3 <- MIDO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs E. Coli CFU
```{r}
BS_ECFU <- MIDO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
```{r}
ECFU_Con <- MIDO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Turbidity
```{r}
ECFU_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Nitrate Concentration
```{r}
ECFU_no3 <- MIDO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

## NORO

@@

#### Biological Vs Conductivity
```{r}
N_BS_Con <- NORO %>%
  ggplot(aes(x=conductivity.uscm, y=biological_score)) +
  geom_point(color = "red") +
   geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Turbidity
```{r}
N_BS_Turb <- NORO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Nitrate Concentration
```{r}
N_BS_no3 <- NORO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs E. Coli CFU
```{r}
N_BS_ECFU <- NORO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
```{r}
N_ECFU_Con <- NORO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Turbidity
```{r}
N_ECFU_Turb <- NORO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black")

@@

#### E. Coli Vs Nitrate Concentration
```{r}
N_ECFU_no3 <- NORO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

## BICO

@@

#### Biological Vs Conductivity
```{r}
B_BS_Con <- BICO %>%
  ggplot(aes(x=conductivity.uscm, y=biological_score)) +
  geom_point(color = "red") +
   geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Turbidity
```{r}
B_BS_Turb <- BICO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Nitrate Concentration
```{r}
B_BS_no3 <- BICO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs E. Coli CFU
```{r}
B_BS_ECFU <- BICO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
```{r}
B_ECFU_Con <- BICO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Turbidity
```{r}
B_ECFU_Turb <- BICO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Nitrate Concentration
```{r}
B_ECFU_no3 <- BICO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

### In summary, exploration of correlation of stream health indicators showed that most interactions were non-significant. However, it is to be noted that the ECFU and turbidity interaction was significantly positive. In further analysis, this may serve as a check to show that correlative analyses are robust enough with data provided. In aquatic microbiology, the positive relationship between ECFU growth and turbidity is generally accepted because the more bacteria that grow in the water column, the cloudier it will be.

>>>>

#### First, we need to load all the libraries we will need.
```{r}
library(dplyr) 
library(here)
library(tidyverse)
library(tidyr)
library(readr)
library(stringr)
library(gridExtra)
```

@@

## MIDO
_This section of regression analysis and plotting deals exclusively with the MIDO data set._

@@

#### Biological Vs Conductivity:
_This plot shows the relationship between biological score as a function of conductivity._
```{r}
BS_Con <- MIDO %>% 
  ggplot(aes(x=(conductivity.uscm), 
             y=(biological_score))) +
  geom_point(color = "red") +
  geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Turbidity:
_This plot shows the relationship between biological score as a function of turbidity._
```{r}
BS_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Nitrate Concentration
_This plot shows the relationship between biological score as a function of nitrate (no3) concentration._
```{r}
BS_no3 <- MIDO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs E. Coli CFU
_This plot shows the relationship between biological score as a function of E. Coli (cfu)._
```{r}
BS_ECFU <- MIDO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Now let's use the gridExtra package to view all the MIDO Biological Indicator outcome relationship graphs together.
```{r}
#Creates a plot with all the MIDO biological score relationships in it.
MIDO_bs_lm <- grid.arrange(BS_Con, BS_ECFU, BS_no3, BS_Turb, ncol = 2)
```
</br>
_Save MIDO Biological Score figures_
```{r}
#Stores the location I want to save the graphs to.
MIDO_bs_lm_location <- here::here("results","MIDO_bs_lm.png")
#Saves graphs to stored location in ,my working directory.
ggsave(MIDO_bs_lm, file = MIDO_bs_lm_location)
```
</br>
### *E. Coli*
</br>
#### E. Coli Vs Conductivity
_This plot shows the relationship between E. coli (cfu) as a function of Conductivity._
```{r}
ECFU_Con <- MIDO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Turbidity
_This plot shows the relationship between E. coli as a function of Turbidity._
```{r}
ECFU_Turb <- MIDO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Nitrate Concentration
_This plot shows the relationship between E. coli (cfu) as a function of nitrate (no3) concentration._
```{r}
ECFU_no3 <- MIDO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Now let's use the gridExtra package to view all the MIDO E. coli CFU outcome relationship graphs together.
```{r}
MIDO_ECFU_lm <- grid.arrange(ECFU_Con, ECFU_no3, ECFU_Turb, ncol = 2)
```
</br>
_Save MIDO ECFU figures_
```{r}
#Stores the location I want to save the graphs to.
MIDO_ECFU_lm_location <- here::here("results","MIDO_ECFU_lm.png")
#Saves graphs to stored location in ,my working directory.
ggsave(MIDO_ECFU_lm, file = MIDO_ECFU_lm_location)
```
</br>
## NORO
_the following linear analysis deals exclusively with the NORO dataset._

@@

#### Biological Vs Conductivity
_This plot shows the relationship between Biological Score as a function of Conductivity._
```{r}
N_BS_Con <- NORO %>%
  ggplot(aes(x=conductivity.uscm, y=biological_score)) +
  geom_point(color = "red") +
   geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Turbidity
_This plot shows the relationship between Biological Score as a function of Turbidity._
```{r}
N_BS_Turb <- NORO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Nitrate Concentration
_This plot shows the relationship between Biological Score as a function of nitrate (no3) Concentration._
```{r}
N_BS_no3 <- NORO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs E. Coli CFU\
_This plot shows the relationship between Biological Score as a function of E. coli (cfu)._
```{r}
N_BS_ECFU <- NORO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

</br>
#### Now let's use the gridExtra package to view all the NORO Biological Indicator outcome relationship graphs together.
```{r}
NORO_bs_lm <- grid.arrange(N_BS_Con, N_BS_ECFU, N_BS_no3, N_BS_Turb, ncol = 2)
```
</br>
_Save NORO Biological Score figures_
```{r}
#Stores the location I want to save the graphs to.
NORO_bs_lm_location <- here::here("results","NORO_bs_lm.png")
#Saves graphs to stored location in ,my working directory.
ggsave(NORO_bs_lm, file = NORO_bs_lm_location)
```
</br>
</br>

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
_This plot shows the relationship between E. Coli (cfu) as a function of Conductivity._
```{r}
N_ECFU_Con <- NORO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Turbidity
_This plot shows the relationship between E. Coli (cfu) as a function of Turbidity._
```{r}
N_ECFU_Turb <- NORO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black")

@@

#### E. Coli Vs Nitrate Concentration
_This plot shows the relationship between E. coli as a function of nitrate (no3) concentration._
```{r}
N_ECFU_no3 <- NORO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Now let's use the gridExtra package to view all the NORO E. coli CFU outcome relationship graphs together.
```{r}
NORO_ECFU_lm <- grid.arrange(N_ECFU_Con, N_ECFU_no3, N_ECFU_Turb, ncol = 2)
```
</br>
_Save NORO ECFU figures_
```{r}
#Stores the location I want to save the graphs to.
NORO_ECFU_lm_location <- here::here("results","NORO_ECFU_lm.png")
#Saves graphs to stored location in ,my working directory.
ggsave(NORO_ECFU_lm, file = NORO_ECFU_lm_location)
```

## BICO:
_The following Linear Regression analysis deals exclusively with the BICO data set._

@@

#### Biological Vs Conductivity
_This plot shows the relationship between biological score as a function of conductivity._
```{r}
B_BS_Con <- BICO %>%
  ggplot(aes(x=conductivity.uscm, y=biological_score)) +
  geom_point(color = "red") +
   geom_smooth(method=lm, color = "red") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Turbidity
_This plot shows the relationship between biological score as a function of turbidity._
```{r}
B_BS_Turb <- BICO %>%
  ggplot(aes(x=turbidity.ntu, y=biological_score)) +
  geom_point(color = "blue") +
   geom_smooth(method=lm, color = "blue") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs Nitrate Concentration
_This plot shows the relationship between biological score as a function of nitrate (no3) concentration._
```{r}
B_BS_no3 <- BICO %>%
  ggplot(aes(x=no3.mgL, y=biological_score)) +
  geom_point(color = "green") +
   geom_smooth(method=lm, color = "green") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Biological Vs E. Coli CFU
_This plot shows the relationship between biological score as a function of E. coli (cfu)._
```{r}
B_BS_ECFU <- BICO %>%
  ggplot(aes(x=e.coli.cfu, y=biological_score)) +
  geom_point(color = "purple") +
   geom_smooth(method=lm, color = "purple") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Now let's use the gridExtra package to view all the BICO Biological Indicator outcome relationship graphs together.
```{r}
BICO_bs_lm <- grid.arrange(B_BS_Con, B_BS_ECFU, B_BS_no3, B_BS_Turb, ncol = 2)
```
</br>
_Save BICO biological score figures_
```{r}
#Stores the location I want to save the graphs to.
BICO_bs_lm_location <- here::here("results","BICO_bs_lm.png")
#Saves graphs to stored location in ,my working directory.
ggsave(BICO_bs_lm, file = BICO_bs_lm_location)
```

### *E. Coli*
</br>
#### E. Coli Vs Conductivity
_This plot shows the relationship between E. coli (cfu) as a function of conductivity._
```{r}
B_ECFU_Con <- BICO %>%
  ggplot(aes(x=conductivity.uscm, y=e.coli.cfu)) +
  geom_point(color = "orange") +
   geom_smooth(method=lm, color = "orange") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Turbidity
_This plot shows the relationship between E. coli (cfu) as a function of turbidity._
```{r}
B_ECFU_Turb <- BICO %>%
  ggplot(aes(x=turbidity.ntu, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### E. Coli Vs Nitrate Concentration
_This plot shows the relationship between E. coli (cfu) as a function of nitrate (no3) concentration._
```{r}
B_ECFU_no3 <- BICO %>%
  ggplot(aes(x=no3.mgL, y=e.coli.cfu)) +
  geom_point(color = "black") +
   geom_smooth(method=lm, color = "black") +
  scale_x_continuous(trans = 'log10') +
  scale_y_continuous(trans = 'log10')

@@

#### Now let's use the gridExtra package to view all the BICO E. coli CFU outcome relationship graphs together.
```{r}
BICO_ECFU_lm <- grid.arrange(B_ECFU_Con, B_ECFU_no3, B_ECFU_Turb, ncol = 2)
```
</br>
_Save BICO ECFU figures_
```{r}
#Stores the location I want to save the graphs to.
BICO_ECFU_lm_location <- here::here("results","BICO_ECFU_lm.png")
#Saves graphs to stored location in ,my working directory.
ggsave(BICO_ECFU_lm, file = BICO_ECFU_lm_location)
```
</br>
### Save the exploratory linear models
#### MIDO
_Biological Score_
```{r}
BS_Con_location <- here::here("data","processed_data","BS_Con.rds")
saveRDS(BS_Con_lm, file = BS_Con_location)

BS_ECFU_location <- here::here("data","processed_data","BS_ECFU.rds")
saveRDS(BS_ECFU_lm, file = BS_ECFU_location)

BS_no3_location <- here::here("data","processed_data","BS_no3.rds")
saveRDS(BS_no3_lm, file = BS_no3_location)

BS_Turb_location <- here::here("data","processed_data","BS_Turb.rds")
saveRDS(BS_Turb_lm, file = BS_Turb_location)
```
_ECFU_
```{r}
ECFU_Con_location <- here::here("data","processed_data","BS_Con.rds")
saveRDS(ECFU_Con_lm, file = ECFU_Con_location)

ECFU_no3_location <- here::here("data","processed_data","BS_no3.rds")
saveRDS(ECFU_no3_lm, file = ECFU_no3_location)

ECFU_Turb_location <- here::here("data","processed_data","BS_Turb.rds")
saveRDS(ECFU_Turb_lm, file = ECFU_Turb_location)
```

#### BICO
_Biological Score_
```{r}
B_BS_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
saveRDS(B_BS_Con_lm, file = B_BS_Con_location)

B_BS_ECFU_location <- here::here("data","processed_data","B_BS_ECFU.rds")
saveRDS(B_BS_ECFU_lm, file = B_BS_ECFU_location)

B_BS_no3_location <- here::here("data","processed_data","B_BS_no3.rds")
saveRDS(B_BS_no3_lm, file = B_BS_no3_location)

B_BS_Turb_location <- here::here("data","processed_data","B_BS_Turb.rds")
saveRDS(B_BS_Turb_lm, file = B_BS_Turb_location)
```
_ECFU_
```{r}
B_ECFU_Con_location <- here::here("data","processed_data","B_BS_Con.rds")
saveRDS(B_ECFU_Con_lm, file = B_ECFU_Con_location)

B_ECFU_no3_location <- here::here("data","processed_data","B_BS_no3.rds")
saveRDS(B_ECFU_no3_lm, file = B_ECFU_no3_location)

B_ECFU_Turb_location <- here::here("data","processed_data","B_BS_Turb.rds")
saveRDS(B_ECFU_Turb_lm, file = B_ECFU_Turb_location)
```

#### NORO
_Biological Score_
```{r}
N_BS_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
saveRDS(N_BS_Con_lm, file = N_BS_Con_location)

N_BS_ECFU_location <- here::here("data","processed_data","N_BS_ECFU.rds")
saveRDS(N_BS_ECFU_lm, file = N_BS_ECFU_location)

N_BS_no3_location <- here::here("data","processed_data","N_BS_no3.rds")
saveRDS(N_BS_no3_lm, file = N_BS_no3_location)

N_BS_Turb_location <- here::here("data","processed_data","N_BS_Turb.rds")
saveRDS(N_BS_Turb_lm, file = N_BS_Turb_location)
```
_ECFU_
```{r}
N_ECFU_Con_location <- here::here("data","processed_data","N_BS_Con.rds")
saveRDS(N_ECFU_Con_lm, file = N_ECFU_Con_location)

N_ECFU_no3_location <- here::here("data","processed_data","N_BS_no3.rds")
saveRDS(N_ECFU_no3_lm, file = N_ECFU_no3_location)

N_ECFU_Turb_location <- here::here("data","processed_data","N_BS_Turb.rds")
saveRDS(N_ECFU_Turb_lm, file = N_ECFU_Turb_location)
```

</br>
### In summary, exploration of correlation of stream health indicators showed that most interactions were non-significant. However, it is to be noted that the ECFU and turbidity interaction was significantly positive. In further analysis, this may serve as a check to show that correlative analyses are robust enough with data provided. In aquatic microbiology, the positive relationship between ECFU growth and turbidity is generally accepted because the more bacteria that grow in the water column, the cloudier it will be.

### 2021-10-29 products/manuscript/Manuscript.Rmd
Project Part 3 - CC

---
title: "Relationship Between Common Stream Health Indicators"
subtitle: ""
author: Carter Coleman
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: ../references.bib
csl: ../apa.csl
---

# Summary/Abstract

@@

@@

## Data aquisition
_The data I am using is the publically avaialable data provided by the Upper Oconee Watershed Network (UOWN). This data is part of a citizen science training effort for routine watershed-scale monitoring of the upper oconee river watershed. Trained freshwater scientists assist in training members of the community to collect both chemical and biological stream characteristics that are useful in describing stream health. All data is compliled and posted in raw form on the UOWN website. Of interest to me are the following variables: WSID (ID number),	biological_score,	conductivity, turbidity,	no3.mgL,	e.coli.cfu,	month,	year,	day. The raw data shows 3,075 observations before cleaning. However, while cleaning I had to eliminate "po4.mgl" and "temperature" due to significant amounts of missing data._

#### Load import packages:
```{r}
library(tidyverse)
library(readxl)
```

#### Lodaing data and providing a summary for the raw data
```{r, warning=FALSE}
data_location <- here::here("data/raw_data/UOWN_data_master_spring2021 (2).xlsx")

@@

glimpse(rawdata)
```

@@

## Exploratory analysis

## Full analysis

# Discussion

>>>>

---
title: "Relationship Between Common Stream Health Indicators"
subtitle: ""
author: Carter Coleman
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
documentclass: article
output:
  bookdown::word_document2: 
    toc: false
    number_sections: true
  bookdown::html_document2: 
    toc: false
bibliography: ../references.bib
csl: ../apa.csl
---
# Summary/Abstract

</br>

@@

</br>

@@

</br>
## Data aquisition
_The data I am using is the publically avaialable data provided by the Upper Oconee Watershed Network (UOWN). This data is part of a citizen science training effort for routine watershed-scale monitoring of the upper oconee river watershed. Trained freshwater scientists assist in training members of the community to collect both chemical and biological stream characteristics that are useful in describing stream health. All data is compliled and posted in raw form on the UOWN website. Of interest to me are the following variables: WSID (ID number),	biological_score,	conductivity, turbidity,	no3.mgL,	e.coli.cfu,	month,	year,	day. The raw data shows 3,075 observations before cleaning. However, while cleaning I had to eliminate "po4.mgl" and "temperature" due to significant amounts of missing data._
</br>
#### Load import packages:
```{r}
library(tidyverse)
library(readxl)
library(here)
library(png)
library(knitr)
```
</br>
#### Lodaing data and providing a summary for the raw data
```{r, warning=FALSE}
data_location <- here::here("data/raw_data/UOWN_data_master_spring2021 (2).xlsx")

@@

glimpse(rawdata)
```
</br>

@@

_Below is the cleaned data produced from our data processing. Data was processed using the processingscript.R file and includes the following steps:_
_1. Load rawdata from excel file. Note the use of the here() package and not absolute paths_
</br>
_2. Designate NA's as actually missing values instead of factor data._
</br>
_3. Select variables of interest. These are variables I think are good indicators of stream health or are important identifying labels. Such variables are listed here: WSID, biological_score, conductivity.uscm, turbidity.ntu, po4.mgL, no3.mgL, pH, temperature.c,	e.coli.cfu, month, year,	and day. This was done using the Select() function._
</br>
_There is 3,075 rows and 12 columns... So, a lot of data. This is because data has been collected for ~20 years. Unfortunately, I am seeing a lot of NA's in the data. Because this is citizen science monitoring data, NA most likely means the data wasn't collected during that event and is a form of human error bias._
</br>
_4. Use NA assessment to see determine which variables to remove from assessment. This was done using the is_na function._
_Several of the columns have more than half of the data missing. This is pretty common for citizen science data, seeing as collection is opportunistic. As such, I decided to remove P04 and temperature, because they are missing a lot of data. Finally, I decided to remove all missing values and see the condition of the data knowing that there are now no missing values. I removed NA data using the na.omit() function_
</br>
_5. Change all numeric columns to be read as numeric instead of character. This sometimes happens when R tries to read in data from an external source, but is an easy fix with the as.numeric function. This is important because my analysis uses linear regression functions and they are coded to only deal with numeric values._
 </br>                          
_6. Separate the data into three separate dataframes based on location ID: "MIDO", "NORO", and "BICO". This is to separate out any bias based on location of the streams within ACC._
</br>
_7. Because the number of rows between each site data frame and between years is not consistent, measuring between sites and through time may be biased. That being said, my original question refers to the correlation between different indicators of stream health, not how they change spatial or temporally. Therefore, these space and time inconsistencies won't affect the analyses._
</br>
_8. Saved cleand data files_
```{r}
#Set the location of the desired dataframes
MIDO_location <- here::here("data","processed_data","MIDO.RDS")
NORO_location <- here::here("data","processed_data","NORO.RDS")
BICO_location <- here::here("data","processed_data","BICO.RDS")

#Load in the Dataframes
MIDO <- readRDS(MIDO_location)
NORO <- readRDS(NORO_location)
BICO <- readRDS(BICO_location)
```
## Exploratory analysis
_The following is an exploration of Stream Health Indicator correlation and comparison. The exploration will be structured as follows:_
_Each dataframe (MIDO, NORO, and BICO) will be analyzed separately. Within each dataframe, there are data explaining biological indicator values, stream chemistry, and fecal indicators over time. The only group with multiple components is stream chemistry. I plan to do a series of correlative analysis showing if there is some sort of relationship between the three indicator types to help inform further analysis via principle component, multivariate analysis, etc. Therefore, the goal of this exploration is basically to summarize how these indicators correlate._
### Below is the Linear Regression analysis from the Exploratory Data Analysis
#### MIDO
_Outcome: Biological Score_
```{r}
#Define data location
MIDO_bs_lm_location <- here::here("results","MIDO_bs_lm.png")
#Load in saved image from results folder
include_graphics(MIDO_bs_lm_location)
```
_Outcome: E. coli (cfu)_
```{r}
#Define data location
MIDO_ecfu_lm_location <- here::here("results","MIDO_ECFU_lm.png")
#Load in the data
include_graphics(MIDO_ecfu_lm_location)
```

#### BICO
_Outcome: Biological Score_
```{r}
#Define data location
BICO_bs_lm_location <- here::here("results","BICO_bs_lm.png")
#Load in saved image from results folder
include_graphics(BICO_bs_lm_location)
```
_Outcome: E. coli (cfu)_
```{r}
#Define data location
BICO_ecfu_lm_location <- here::here("results","BICO_ECFU_lm.png")
#Load in the data
include_graphics(BICO_ecfu_lm_location)
```

#### NORO
_Outcome: Biological Score_
```{r}
#Define data location
NORO_bs_lm_location <- here::here("results","NORO_bs_lm.png")
#Load in saved image from results folder
include_graphics(NORO_bs_lm_location)
```
_Outcome: E. coli (cfu)_
```{r}
#Define data location
NORO_ecfu_lm_location <- here::here("results","NORO_ECFU_lm.png")
#Load in the data
include_graphics(NORO_ecfu_lm_location)
```
## Full analysis
_At this point,the main take-aways from data analysis thus far is that correlation of stream health indicators showed that most interactions were non-significant, with the exception of E. coli (cfu) and turbidity. Additionally, the data set is most likely not skewed by influential outliers._
</br>
_Moving forward, I want to focus in on just the MIDO dataset. As you can tell, working with many variables in three different data sets can be quite cumbersome to read through. For ease of interpretation, the MIDO data set will serve as the data frame where I test how robust the linear fit models actually are for the data. MIDO is the only data set large enough that when we split the data set into a "test" and "train" data set, I actually have confidence in the resulting model fit analysis. For comparison, MIDO has an n = 88; NORO has an n = 35; and BICO has an n = 23. The only thing we are missing by dropping the other sites is assessing site-by-site comparison, which I am choosing not to focus on anyway._
</br>
_The results indicated that only the relationship between E. Coli (cfu) and Conductivity were significant for both the train data (p = 1.374747e-05) and test data (p = 1.89214e-06). However, both models had high RMSE values, indicating that the models did not fit the data well and could not be used to predict E. coli (cfu) from conductivity, despite a significant relationship._
</br>
# Discussion
